Realizability-Preserving DG-IMEX Method for the Two-Moment Model of Fermion Transport \tnoteref{support}

### 2018-02-03 Documents/M1/realizableFermionicM1/sections/DGIMEXscheme.tex
Additions to DG-IMEX scheme

Next, we establish conditions for which \eqref{eq:realizableGamma} holds.  
To simplify the notation, we temporarily drop the dimension index $i$, setting $x^{i}=x$, $K^{i}=K$, $\vect{\Gamma}^{i}=\vect{\Gamma}$, etc.  
We let $\hat{Q}_{N}$ denote the $N$-point \emph{Gauss-Lobatto} quadrature rule on the interval $K=(x_{\Low},x_{\Hgh})$, with points
\begin{equation}
  \hat{S}=\left\{x_{\Low}=\hat{x}_{1},\cdots,\hat{x}_{N}=x_{\Hgh}\right\}, 
  \label{eq:quadraturePointsGL}
\end{equation}
and weights $\hat{w}_{q} \in (0,1]$, normalized so that $\sum_{q=1}^{N} \hat{w}_{q} = 1$.  
(The hat is used to denote the Gauss-Lobatto rule, which includes the endpoints of the interval $K$.)
This quadrature integrates polynomials in $x \in \bbR$ with degree $\le2N-3$ exactly.

>>>>

Next, we establish conditions for which \eqref{eq:realizableGamma} holds.  
To simplify the notation, we temporarily drop the dimension index $i$, setting $x^{i}=x$, $K^{i}=K$, $\vect{\Gamma}^{i}=\vect{\Gamma}$, etc.  
We let $\hat{Q}_{N}$ denote the $N$-point \emph{Gauss-Lobatto (GL)} quadrature rule on the interval $K=(x_{\Low},x_{\Hgh})$, with points
\begin{equation}
  \hat{S}=\left\{x_{\Low}=\hat{x}_{1},\cdots,\hat{x}_{N}=x_{\Hgh}\right\}, 
  \label{eq:quadraturePointsGL}
\end{equation}
and weights $\hat{w}_{q} \in (0,1]$, normalized so that $\sum_{q=1}^{N} \hat{w}_{q} = 1$.  
(The hat is used to denote the GL rule, which includes the endpoints of the interval $K$.)
This quadrature integrates polynomials in $x \in \bbR$ with degree $\le2N-3$ exactly.  
Then, if $\vect{\cM}_{h}$ is represented by such polynomials we have
\begin{equation}
  \int_{K} \bcM_{h}\,dx = \hat{Q}_{N}[\bcM_{h}] \equiv
  |K| \sum_{q=1}^{N} \hat{w}_{q}\,\hat{\bcM}_{q}
  \label{eq:quadratureRuleGL}
\end{equation}
where $\hat{\bcM}_{q} := \bcM_{h}(\hat{x}_q,\tilde{\vect{x}})$.  
In each element, $\hat{\vect{\cM}}_{1}=\vect{\cM}_{\Low}^{+}$ and $\hat{\vect{\cM}}_{N}=\vect{\cM}_{\Hgh}^{-}$.

With the GL quadrature rule in \eqref{eq:quadratureRuleGL} (note $\hat{w}_{1}=\hat{w}_{N}$), we can write $\vect{\Gamma}$ as the convex combination
\begin{equation}
  \vect{\Gamma}\big[\bcM_{h}\big]
  =\sum_{q=2}^{N-1}\hat{w}_{q}\,\hat{\vect{\cM}}_{q}
  +2\,\hat{w}_{N}\,\Phi(\vect{\cM}_{\Low}^{-},\vect{\cM}_{\Low}^{+},\vect{\cM}_{\Hgh}^{-},\vect{\cM}_{\Hgh}^{+}),
  \label{eq:realizableGammaConvex}
\end{equation}
where we have defined
\begin{align}
  &\Phi(\vect{\cM}_{\Low}^{-},\vect{\cM}_{\Low}^{+},\vect{\cM}_{\Hgh}^{-},\vect{\cM}_{\Hgh}^{+}) \nonumber \\
  &\hspace{12pt}
  =\f{1}{2}\big(\,\vect{\cM}_{\Low}^{+}+\lambda\,\mathscr{F}(\vect{\cM}_{\Low}^{-},\vect{\cM}_{\Low}^{+})\,\big)
  +\f{1}{2}\big(\,\vect{\cM}_{\Hgh}^{-}-\lambda\,\mathscr{F}(\vect{\cM}_{\Hgh}^{-},\vect{\cM}_{\Hgh}^{+})\,\big),
\end{align}
with $\lambda=\hat{c}\,\dt/(s\,\hat{w}_{N}\,|K|)$.  
The following Lemma establishes sufficient conditions for realizability of $\vect{\Gamma}$, and hence $\vect{\cM}_{\bK}^{*}$.  
\begin{lemma}
  Assume that $\hat{\vect{\cM}}_{q}\in\cR$ for all $q=1,\ldots,N$ and all $\bK\in\mathscr{T}$.  
  Let the time step $\dt$ be chosen so that $\lambda\le1$.  
  Let the numerical flux be given by the Lax-Friedrichs flux in Eq.~\eqref{eq:fluxFunctionLF} with $\alpha=1$.  
  Then $\vect{\Gamma}[\bcM_{h}]\in\cR$.  
\end{lemma}
\begin{proof}
  In Eq.~\eqref{eq:realizableGammaConvex}, $\vect{\Gamma}[\vect{\cM}_{h}]$ is expressed as a convex combination.  
  By assumption, $\hat{\vect{\cM}}_{q}\in\cR$ ($q=2,\ldots,N-1$).  
  It remains to show $\Phi(\vect{\cM}_{\Low}^{-},\vect{\cM}_{\Low}^{+},\vect{\cM}_{\Hgh}^{-},\vect{\cM}_{\Hgh}^{+})\in\cR$.  
  Using the Lax-Friedrichs flux with $\alpha=1$, it is straightforward to show that
  \begin{align}
    &\Phi(\vect{\cM}_{\Low}^{-},\vect{\cM}_{\Low}^{+},\vect{\cM}_{\Hgh}^{-},\vect{\cM}_{\Hgh}^{+})
    =(1-\lambda)\,\f{1}{2}\,\big(\,\vect{\cM}_{\Low}^{+}+\vect{\cM}_{\Hgh}^{-}\,\big) \nonumber \\
    &\hspace{12pt}
    +\f{1}{2}\,\lambda\,\big(\,\Phi^{+}(\vect{\cM}_{\Low}^{-})+\Phi^{-}(\vect{\cM}_{\Hgh}^{-})\,\big)
    +\f{1}{2}\,\lambda\,\big(\,\Phi^{+}(\vect{\cM}_{\Low}^{+})+\Phi^{-}(\vect{\cM}_{\Hgh}^{+})\,\big),
    \label{eq:phiConvex}
  \end{align}
  where $\Phi^{\pm}(\vect{\cM})=\f{1}{2}\,\big(\vect{\cM}\pm\vect{e}\cdot\vect{\cF}(\vect{\cM})\big)$.  
  Since $\lambda\le1$, $\Phi$ is expressed as a convex combination of three terms.  
  By assumption, $\vect{\cM}_{\Low}^{-},\vect{\cM}_{\Low}^{+},\vect{\cM}_{\Hgh}^{-},\vect{\cM}_{\Hgh}^{+}\in\cR$, which immediately implies realizability of the first term on the right-hand side of Eq.~\eqref{eq:phiConvex}.  
  Realizability of the second and third terms follows by invoking Lemma~\ref{lem:explicitStep}.  
  This completes the proof.  
\end{proof}
